---
title: "Palmer Penguins Dashboard"
author: "Data Analytics Team"
format:
  dashboard:
    theme: cosmo
    nav-buttons:
      - icon: github
        href: https://github.com/allisonhorst/palmerpenguins
    logo: https://allisonhorst.github.io/palmerpenguins/logo.png
    scrolling: false
execute:
  warning: false
  message: false
---

```{r}
#| label: setup

library(tidyverse)
library(palmerpenguins)
library(plotly)
library(DT)
library(bslib)
library(bsicons)

# Remove NAs for cleaner dashboard
penguins_clean <- penguins %>% drop_na()

# Color palette
species_colors <- c("Adelie" = "#FF8C00",
                    "Chinstrap" = "#A034F0",
                    "Gentoo" = "#159090")

# Calculate metrics
total_penguins <- nrow(penguins_clean)
avg_body_mass <- mean(penguins_clean$body_mass_g)
num_islands <- n_distinct(penguins_clean$island)
study_years <- paste(min(penguins_clean$year), "-", max(penguins_clean$year))
```

# Overview {orientation="columns"}

## Column {width="25%"}

### Row {height="25%"}

```{r}
#| label: value-box-total
#| content: valuebox
#| title: "Total Penguins"

list(
  icon = "clipboard-data",
  color = "primary",
  value = total_penguins
)
```

### Row {height="25%"}

```{r}
#| label: value-box-mass
#| content: valuebox
#| title: "Avg Body Mass"

list(
  icon = "arrow-up-circle",
  color = "success",
  value = paste0(round(avg_body_mass), " g")
)
```

### Row {height="25%"}

```{r}
#| label: value-box-islands
#| content: valuebox
#| title: "Islands Studied"

list(
  icon = "geo-alt",
  color = "info",
  value = num_islands
)
```

### Row {height="25%"}

```{r}
#| label: value-box-years
#| content: valuebox
#| title: "Study Period"

list(
  icon = "calendar-range",
  color = "warning",
  value = study_years
)
```

## Column {width="75%"}

### Row {height="50%"}

```{r}
#| label: species-scatter
#| title: "Body Mass vs Flipper Length by Species"

p <- penguins_clean %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g,
             color = species, text = paste0(
               "Species: ", species, "<br>",
               "Flipper: ", flipper_length_mm, " mm<br>",
               "Mass: ", body_mass_g, " g<br>",
               "Island: ", island, "<br>",
               "Sex: ", sex
             ))) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = species_colors) +
  labs(
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species"
  ) +
  theme_minimal(base_size = 13)

ggplotly(p, tooltip = "text") %>%
  layout(
    hovermode = "closest",
    legend = list(
      orientation = "h",
      y = -0.15,
      x = 0.5,
      xanchor = "center"
    )
  )
```

### Row {height="50%"}

```{r}
#| label: bill-dimensions
#| title: "Bill Dimensions Comparison"

p <- penguins_clean %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm,
             color = species, text = paste0(
               "Species: ", species, "<br>",
               "Bill Length: ", bill_length_mm, " mm<br>",
               "Bill Depth: ", bill_depth_mm, " mm<br>",
               "Island: ", island
             ))) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = species_colors) +
  labs(
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)",
    color = "Species"
  ) +
  theme_minimal(base_size = 13)

ggplotly(p, tooltip = "text") %>%
  layout(
    hovermode = "closest",
    legend = list(
      orientation = "h",
      y = -0.15,
      x = 0.5,
      xanchor = "center"
    )
  )
```

# Species Analysis {orientation="rows"}

## Row {height="40%"}

### Column {width="33%"}

```{r}
#| label: adelie-box
#| title: "Adelie Penguins"

adelie_count <- penguins_clean %>%
  filter(species == "Adelie") %>%
  nrow()

adelie_avg_mass <- penguins_clean %>%
  filter(species == "Adelie") %>%
  summarise(avg = mean(body_mass_g)) %>%
  pull(avg)

adelie_islands <- penguins_clean %>%
  filter(species == "Adelie") %>%
  distinct(island) %>%
  pull(island) %>%
  paste(collapse = ", ")

value_box(
  title = "Adelie Overview",
  value = adelie_count,
  showcase = bs_icon("clipboard-data"),
  theme = "primary",
  p(paste0("Avg Mass: ", round(adelie_avg_mass), " g")),
  p(paste0("Islands: ", adelie_islands))
)
```

### Column {width="33%"}

```{r}
#| label: chinstrap-box
#| title: "Chinstrap Penguins"

chinstrap_count <- penguins_clean %>%
  filter(species == "Chinstrap") %>%
  nrow()

chinstrap_avg_mass <- penguins_clean %>%
  filter(species == "Chinstrap") %>%
  summarise(avg = mean(body_mass_g)) %>%
  pull(avg)

chinstrap_islands <- penguins_clean %>%
  filter(species == "Chinstrap") %>%
  distinct(island) %>%
  pull(island) %>%
  paste(collapse = ", ")

value_box(
  title = "Chinstrap Overview",
  value = chinstrap_count,
  showcase = bs_icon("clipboard-data"),
  theme = "warning",
  p(paste0("Avg Mass: ", round(chinstrap_avg_mass), " g")),
  p(paste0("Islands: ", chinstrap_islands))
)
```

### Column {width="33%"}

```{r}
#| label: gentoo-box
#| title: "Gentoo Penguins"

gentoo_count <- penguins_clean %>%
  filter(species == "Gentoo") %>%
  nrow()

gentoo_avg_mass <- penguins_clean %>%
  filter(species == "Gentoo") %>%
  summarise(avg = mean(body_mass_g)) %>%
  pull(avg)

gentoo_islands <- penguins_clean %>%
  filter(species == "Gentoo") %>%
  distinct(island) %>%
  pull(island) %>%
  paste(collapse = ", ")

value_box(
  title = "Gentoo Overview",
  value = gentoo_count,
  showcase = bs_icon("clipboard-data"),
  theme = "success",
  p(paste0("Avg Mass: ", round(gentoo_avg_mass), " g")),
  p(paste0("Islands: ", gentoo_islands))
)
```

## Row {height="60%"}

### Column {width="50%"}

```{r}
#| label: species-distribution
#| title: "Species Distribution by Island and Sex"

species_dist <- penguins_clean %>%
  count(species, island, sex) %>%
  ggplot(aes(x = island, y = n, fill = species)) +
  geom_col(position = "dodge") +
  facet_wrap(~sex) +
  scale_fill_manual(values = species_colors) +
  labs(
    x = "Island",
    y = "Count",
    fill = "Species"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

ggplotly(species_dist) %>%
  layout(
    legend = list(
      orientation = "h",
      y = -0.2,
      x = 0.5,
      xanchor = "center"
    )
  )
```

### Column {width="50%"}

```{r}
#| label: mass-distribution
#| title: "Body Mass Distribution by Species"

p <- penguins_clean %>%
  ggplot(aes(x = species, y = body_mass_g, fill = species,
             text = paste0(
               "Species: ", species, "<br>",
               "Mass: ", body_mass_g, " g"
             ))) +
  geom_violin(alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 1) +
  scale_fill_manual(values = species_colors) +
  labs(
    x = "Species",
    y = "Body Mass (g)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")

ggplotly(p, tooltip = "text")
```

# Measurements {orientation="columns"}

## Column {width="40%"}

### Row {height="33%"}

```{r}
#| label: bill-length-box
#| title: "Bill Length by Species"

p <- penguins_clean %>%
  ggplot(aes(x = species, y = bill_length_mm, fill = species)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = species_colors) +
  labs(
    x = "Species",
    y = "Bill Length (mm)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

ggplotly(p) %>%
  layout(showlegend = FALSE)
```

### Row {height="33%"}

```{r}
#| label: bill-depth-box
#| title: "Bill Depth by Species"

p <- penguins_clean %>%
  ggplot(aes(x = species, y = bill_depth_mm, fill = species)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = species_colors) +
  labs(
    x = "Species",
    y = "Bill Depth (mm)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

ggplotly(p) %>%
  layout(showlegend = FALSE)
```

### Row {height="33%"}

```{r}
#| label: flipper-box
#| title: "Flipper Length by Species"

p <- penguins_clean %>%
  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = species_colors) +
  labs(
    x = "Species",
    y = "Flipper Length (mm)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

ggplotly(p) %>%
  layout(showlegend = FALSE)
```

## Column {width="60%"}

### Row {height="50%"}

```{r}
#| label: measurement-heatmap
#| title: "Correlation Heatmap of Measurements"

# Calculate correlations
cor_data <- penguins_clean %>%
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>%
  cor() %>%
  round(2)

# Create heatmap
plot_ly(
  x = colnames(cor_data),
  y = rownames(cor_data),
  z = cor_data,
  type = "heatmap",
  colors = colorRamp(c("#3B4CC0", "#F7F7F7", "#B40426")),
  text = cor_data,
  hovertemplate = "%{y} vs %{x}<br>Correlation: %{z}<extra></extra>"
) %>%
  layout(
    xaxis = list(title = "", tickangle = -45),
    yaxis = list(title = "")
  )
```

### Row {height="50%"}

```{r}
#| label: yearly-trends
#| title: "Average Body Mass Over Years"

yearly_data <- penguins_clean %>%
  group_by(year, species) %>%
  summarise(
    avg_mass = mean(body_mass_g),
    se = sd(body_mass_g) / sqrt(n()),
    .groups = "drop"
  )

plot_ly(yearly_data, x = ~year, y = ~avg_mass,
        color = ~species, colors = species_colors,
        type = "scatter", mode = "lines+markers",
        error_y = list(array = ~se, color = "#777"),
        hovertemplate = paste0(
          "Year: %{x}<br>",
          "Avg Mass: %{y:.0f} g<br>",
          "<extra>%{fullData.name}</extra>"
        )) %>%
  layout(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Average Body Mass (g)"),
    hovermode = "closest",
    legend = list(
      orientation = "h",
      y = -0.2,
      x = 0.5,
      xanchor = "center"
    )
  )
```

# Data Table {orientation="fill"}

```{r}
#| label: data-table
#| title: "Complete Penguin Dataset"

penguins_clean %>%
  select(species, island, bill_length_mm, bill_depth_mm,
         flipper_length_mm, body_mass_g, sex, year) %>%
  datatable(
    options = list(
      pageLength = 20,
      scrollX = TRUE,
      searchHighlight = TRUE,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel')
    ),
    filter = 'top',
    class = 'cell-border stripe hover',
    colnames = c(
      "Species", "Island", "Bill Length (mm)", "Bill Depth (mm)",
      "Flipper Length (mm)", "Body Mass (g)", "Sex", "Year"
    ),
    caption = "Interactive table: Click column headers to sort, use filters to search"
  ) %>%
  formatRound(columns = c("bill_length_mm", "bill_depth_mm",
                          "flipper_length_mm"), digits = 1) %>%
  formatRound(columns = "body_mass_g", digits = 0) %>%
  formatStyle(
    "species",
    backgroundColor = styleEqual(
      c("Adelie", "Chinstrap", "Gentoo"),
      c("#FFE5CC", "#E5CCFF", "#CCF2F2")
    )
  )
```
