---
title: "Palmer Penguins Analysis"
subtitle: "Exploring Antarctic Penguin Populations"
author: "Data Science Team"
date: today
format:
  revealjs:
    theme: [default, custom.scss]
    slide-number: true
    chalkboard: true
    preview-links: auto
    logo: https://allisonhorst.github.io/palmerpenguins/reference/figures/palmerpenguins.png
    footer: "Palmer Penguins | 2026"
    transition: slide
    background-transition: fade
    highlight-style: github
    code-line-numbers: false
execute:
  echo: true
  warning: false
  message: false
---

## Overview {background-color="#2C5F77"}

::: {.incremental}
- **Dataset**: Palmer Penguins
- **Species**: 3 (Adelie, Chinstrap, Gentoo)
- **Location**: Palmer Archipelago, Antarctica
- **Years**: 2007-2009
- **Measurements**: 344 penguins
:::

::: {.notes}
This presentation covers our analysis of the Palmer Penguins dataset.
Key points to emphasize: diversity of species and comprehensive measurements.
:::

---

## Research Questions

::: {.columns}
::: {.column width="50%"}
### Morphology
- Species differences
- Sexual dimorphism
- Measurement correlations
:::

::: {.column width="50%"}
### Distribution
- Island populations
- Temporal trends
- Geographic patterns
:::
:::

::: {.fragment}
::: {.callout-note}
All analyses focus on understanding the ecological patterns across three distinct penguin species.
:::
:::

---

## Data Setup {auto-animate="true"}

```{r}
library(tidyverse)
library(palmerpenguins)
```

::: {.notes}
We'll start by loading our required packages.
:::

---

## Data Setup {auto-animate="true"}

```{r}
library(tidyverse)
library(palmerpenguins)

# Color palette for species
species_colors <- c(
  "Adelie" = "#FF8C00",
  "Chinstrap" = "#A034F0",
  "Gentoo" = "#159090"
)
```

---

## Dataset Structure

```{r}
glimpse(penguins)
```

---

## Species Distribution {.smaller}

::: {.panel-tabset}

### Plot

```{r}
#| echo: false
#| fig-width: 9
#| fig-height: 5

penguins %>%
  drop_na(species, island) %>%
  count(species, island) %>%
  ggplot(aes(x = island, y = n, fill = species)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 5) +
  scale_fill_manual(values = species_colors) +
  labs(
    title = "Penguin Distribution Across Islands",
    x = "Island",
    y = "Number of Penguins",
    fill = "Species"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "top",
    plot.title.position = "plot"
  )
```

### Code

```{r}
#| eval: false
#| code-line-numbers: "|2|3|4|5-7"

penguins %>%
  drop_na(species, island) %>%
  count(species, island) %>%
  ggplot(aes(x = island, y = n, fill = species)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 5)
```

### Key Findings

- **Biscoe**: Primarily Gentoo, some Adelie
- **Dream**: Mix of Adelie and Chinstrap
- **Torgersen**: Exclusively Adelie
- Gentoo penguins are the most numerous overall

:::

::: {.notes}
Notice how each island has a unique composition of species.
Torgersen has only Adelie, making it ecologically distinct.
:::

---

## Body Mass Distribution {background-image="https://allisonhorst.github.io/palmerpenguins/reference/figures/lter_penguins.png" background-opacity="0.1"}

::: {.columns}
::: {.column width="60%"}
```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5.5

penguins %>%
  drop_na(body_mass_g, species) %>%
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_violin(alpha = 0.6, draw_quantiles = c(0.5)) +
  geom_jitter(width = 0.2, alpha = 0.2) +
  scale_fill_manual(values = species_colors) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Body Mass by Species",
    x = "Species",
    y = "Body Mass (g)"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```
:::

::: {.column width="40%"}
### Statistics

```{r}
#| echo: false

penguins %>%
  drop_na(body_mass_g) %>%
  group_by(species) %>%
  summarise(
    Mean = round(mean(body_mass_g)),
    SD = round(sd(body_mass_g)),
    N = n()
  ) %>%
  knitr::kable()
```

::: {.fragment}
Gentoo penguins are significantly larger!
:::
:::
:::

---

## Key Relationship {.smaller}

### Flipper Length vs Body Mass

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 5

penguins %>%
  drop_na(flipper_length_mm, body_mass_g, species) %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.5) +
  scale_color_manual(values = species_colors) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 14)
  )
```

::: {.fragment}
Strong positive correlation across all species (R² > 0.75)
:::

---

## Bill Dimensions

::: {.columns}
::: {.column width="50%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 6

penguins %>%
  drop_na(bill_length_mm, bill_depth_mm, species) %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm,
             color = species)) +
  geom_point(size = 3, alpha = 0.7) +
  stat_ellipse(level = 0.95, linewidth = 1.2) +
  scale_color_manual(values = species_colors) +
  labs(
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)",
    color = "Species"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")
```
:::

::: {.column width="50%"}
::: {.incremental}
### Observations

- Clear species separation
- Adelie: shorter, deeper bills
- Chinstrap: longer, narrower bills
- Gentoo: intermediate dimensions
- Minimal overlap between species
:::
:::
:::

---

## Sexual Dimorphism {.scrollable}

```{r}
#| echo: false
#| fig-width: 11
#| fig-height: 6

penguins %>%
  drop_na(body_mass_g, sex, species) %>%
  ggplot(aes(x = species, y = body_mass_g, fill = sex)) +
  geom_boxplot(alpha = 0.7, width = 0.6) +
  scale_fill_manual(values = c("male" = "#1E88E5", "female" = "#E91E63")) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Body Mass by Species and Sex",
    x = "Species",
    y = "Body Mass (g)",
    fill = "Sex"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 14)
  )
```

::: {.fragment}
Males consistently heavier than females across all species
:::

---

## Statistical Summary {.smaller}

```{r}
#| echo: false

penguins %>%
  drop_na(body_mass_g, sex, species) %>%
  group_by(species, sex) %>%
  summarise(
    N = n(),
    `Mean Mass (g)` = round(mean(body_mass_g)),
    `SD (g)` = round(sd(body_mass_g)),
    .groups = "drop"
  ) %>%
  knitr::kable()
```

::: {.fragment}
::: {.callout-important}
Sex differences range from 500-900g (12-18% of body mass)
:::
:::

---

## Interactive Exploration {.smaller}

```{r}
#| echo: false

library(plotly)

p <- penguins %>%
  drop_na() %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g,
             color = species, shape = sex)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_manual(values = species_colors) +
  labs(
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species",
    shape = "Sex"
  ) +
  theme_minimal(base_size = 12)

ggplotly(p, height = 450)
```

::: {.notes}
Hover over points to see details. Notice the clustering by species and sex.
:::

---

## Key Findings {background-color="#2C5F77"}

::: {.incremental style="font-size: 0.9em;"}
1. **Species Differentiation**: Clear morphological distinctions across all measurements

2. **Size Hierarchy**: Gentoo > Chinstrap > Adelie in most measurements

3. **Sexual Dimorphism**: Consistent male-larger pattern (12-18% difference)

4. **Strong Correlations**: Flipper length and body mass highly correlated (R² > 0.75)

5. **Bill Specialization**: Species-specific bill shapes suggest dietary differences

6. **Geographic Separation**: Island distributions vary by species
:::

---

## Implications {.smaller}

::: {.columns}
::: {.column width="50%"}
### Ecological

- Niche partitioning
- Resource competition
- Habitat preferences
- Breeding patterns
:::

::: {.column width="50%"}
### Research

- Climate change impacts
- Population monitoring
- Conservation strategies
- Long-term trends
:::
:::

::: {.fragment}
::: {.callout-tip}
## Next Steps
Investigate temporal trends and environmental correlates of body size variation
:::
:::

---

## Code Availability

::: {.columns}
::: {.column width="50%"}
### R Package
```r
install.packages("palmerpenguins")
library(palmerpenguins)
data(penguins)
```
:::

::: {.column width="50%"}
### Documentation
- [Package Website](https://allisonhorst.github.io/palmerpenguins/)
- [CRAN Page](https://cran.r-project.org/package=palmerpenguins)
- [GitHub Repo](https://github.com/allisonhorst/palmerpenguins)
:::
:::

---

## Acknowledgments {background-color="#2C5F77"}

::: {style="font-size: 0.9em;"}
**Data Collection**

- Dr. Kristen Gorman
- Palmer Station, Antarctica LTER
- Long Term Ecological Research Network

**Data Package**

- Allison Horst
- Alison Hill
- Kristen Gorman

**Artwork**

- Allison Horst (@allison_horst)
:::

---

## Thank You! {background-color="#2C5F77" background-image="https://allisonhorst.github.io/palmerpenguins/reference/figures/lter_penguins.png" background-opacity="0.2"}

::: {style="font-size: 1.5em; text-align: center; margin-top: 150px;"}
Questions?
:::

::: {style="text-align: center; margin-top: 50px;"}
[allisonhorst.github.io/palmerpenguins](https://allisonhorst.github.io/palmerpenguins/)

📧 contact@example.com
🐦 @dataanalyst
:::
