---
title: "Analysis Report"
author: "Author Name"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    theme: cosmo
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

# Load required packages
library(tidyverse)
library(gt)
library(scales)

# Set global ggplot theme
theme_set(theme_minimal(base_size = 12))

# Set global chunk options
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.width = 8,
  fig.asp = 0.618,
  out.width = "100%"
)
```

## Executive Summary

[Brief overview of key findings and recommendations]

## Introduction

### Background

[Context and motivation for the analysis]

### Objectives

[Specific questions or goals]

### Data Sources

[Description of data used]

```{r}
#| label: data-load
#| include: false

# Load your data
# data <- read_csv("data/dataset.csv")

# For demonstration, using built-in dataset
data <- mtcars %>%
  rownames_to_column(var = "model") %>%
  as_tibble()
```

## Data Overview

### Summary Statistics

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of key variables"

data %>%
  select(mpg, hp, wt, qsec) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarise(
    N = n(),
    Mean = mean(Value),
    SD = sd(Value),
    Min = min(Value),
    Max = max(Value)
  ) %>%
  gt() %>%
  fmt_number(columns = c(Mean, SD, Min, Max), decimals = 2) %>%
  tab_header(title = "Descriptive Statistics") %>%
  cols_label(
    Variable = "Variable",
    N = "N",
    Mean = "Mean",
    SD = "SD",
    Min = "Min",
    Max = "Max"
  )
```

The data contains `r nrow(data)` observations across `r ncol(data) - 1` variables. See @tbl-summary for detailed statistics.

### Data Quality

```{r}
#| label: data-quality

# Check for missing values
missing_data <- data %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing")

total_missing <- sum(missing_data$Missing)
```

The dataset has `r total_missing` missing values across all variables.

## Analysis

### Distribution Analysis

```{r}
#| label: fig-distribution
#| fig-cap: "Distribution of fuel efficiency (MPG)"
#| fig-width: 8
#| fig-height: 5

ggplot(data, aes(x = mpg)) +
  geom_histogram(binwidth = 2, fill = "#2C3E50", alpha = 0.8) +
  geom_density(aes(y = after_stat(count) * 2),
               color = "#E74C3C", linewidth = 1) +
  labs(
    title = "Distribution of Miles Per Gallon",
    x = "Miles Per Gallon (MPG)",
    y = "Count"
  ) +
  theme_minimal()
```

@fig-distribution shows the distribution of fuel efficiency across vehicles.

### Relationship Analysis

```{r}
#| label: fig-scatter
#| fig-cap: "Relationship between weight and fuel efficiency"
#| fig-width: 8
#| fig-height: 6

ggplot(data, aes(x = wt, y = mpg)) +
  geom_point(aes(size = hp, color = as.factor(cyl)), alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "#2C3E50") +
  scale_color_brewer(palette = "Set1", name = "Cylinders") +
  scale_size_continuous(name = "Horsepower") +
  labs(
    title = "Weight vs Fuel Efficiency",
    subtitle = "Colored by number of cylinders, sized by horsepower",
    x = "Weight (1000 lbs)",
    y = "Miles Per Gallon (MPG)"
  ) +
  theme_minimal()
```

As shown in @fig-scatter, there is a strong negative relationship between vehicle weight and fuel efficiency.

### Comparative Analysis

```{r}
#| label: fig-comparison
#| fig-cap: "Fuel efficiency by number of cylinders"
#| fig-width: 8
#| fig-height: 5

data %>%
  mutate(cyl = as.factor(cyl)) %>%
  ggplot(aes(x = cyl, y = mpg, fill = cyl)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.2, alpha = 0.3, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Fuel Efficiency by Engine Configuration",
    x = "Number of Cylinders",
    y = "Miles Per Gallon (MPG)"
  ) +
  theme_minimal()
```

@fig-comparison reveals clear differences in fuel efficiency across engine configurations.

### Statistical Testing

```{r}
#| label: statistical-test

# Perform statistical test
model <- lm(mpg ~ wt + hp, data = data)
r_squared <- summary(model)$r.squared
```

A linear regression model explains `r percent(r_squared, accuracy = 0.1)` of the variance in fuel efficiency.

```{r}
#| label: tbl-regression
#| tbl-cap: "Regression model results"

broom::tidy(model) %>%
  gt() %>%
  fmt_number(columns = c(estimate, std.error, statistic, p.value), decimals = 3) %>%
  tab_header(title = "Linear Regression Results") %>%
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    std.error = "Std. Error",
    statistic = "t-statistic",
    p.value = "p-value"
  )
```

## Key Findings

::: {.callout-note}
## Summary of Results

1. **Finding 1**: [Description]
2. **Finding 2**: [Description]
3. **Finding 3**: [Description]
:::

## Recommendations

::: {.callout-tip}
## Actionable Recommendations

1. **Recommendation 1**: [Action item]
2. **Recommendation 2**: [Action item]
3. **Recommendation 3**: [Action item]
:::

## Limitations

::: {.callout-warning}
## Considerations

- **Limitation 1**: [Description]
- **Limitation 2**: [Description]
:::

## Conclusion

[Summary of analysis and next steps]

## Appendix

### Session Information

```{r}
#| label: session-info
#| echo: true

sessionInfo()
```

### Data Dictionary

```{r}
#| label: tbl-dictionary
#| tbl-cap: "Variable definitions"

tibble(
  Variable = c("mpg", "cyl", "hp", "wt", "qsec"),
  Description = c(
    "Miles per gallon",
    "Number of cylinders",
    "Horsepower",
    "Weight (1000 lbs)",
    "Quarter mile time (seconds)"
  ),
  Type = c("Numeric", "Integer", "Numeric", "Numeric", "Numeric")
) %>%
  gt() %>%
  tab_header(title = "Data Dictionary")
```
